\documentclass{article}
\usepackage{termcal}
\usepackage{amsmath,amsthm,amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{geometry}
\geometry{
  top=1in,
  inner=1in,
  outer=1in,
  bottom=1in,
  headheight=10ex,
  headsep=7ex,
  footskip=10ex
}
% Few useful commands (our classes always meet either on Monday and Wednesday
% or on Tuesday and Thursday)

\newcommand{\MTWFClass}{%
\calday[Monday]{\classday} % Monday
\calday[Tuesday]{\classday} % Tuesday
\calday[Wednesday]{\classday} % Wednesday
\skipday % Thursday (no class)
\calday[Friday]{\classday} % Friday
\skipday\skipday % weekend (no class)
}

\newcommand{\TRClass}{%
\skipday % Monday (no class)
\calday[Tuesday]{\classday} % Tuesday
\skipday % Wednesday (no class)
\calday[Thursday]{\classday} % Thursday
\skipday % Friday
\skipday\skipday % weekend (no class)
}

\newcommand{\Holiday}[2]{%
\options{#1}{\noclassday}
\caltext{#1}{#2}
}

\newcommand{\reading}[1]{\textcolor{OliveGreen}{\textbf{#1}}}
\newcommand{\exam}[1]{\textcolor{RoyalBlue}{\textbf{#1}}}
\newcommand{\hw}[1]{\textcolor{Bittersweet}{\textbf{#1}}}
\newcommand{\noclass}[1]{\textcolor{Gray}{\textbf{#1}}}


\begin{document}
\begin{center}
   \textsc{\large MATH 271, Calendar}\\
   \textsc{Fall 2021}
\vspace{.5cm}

\flushleft\textbf{Color coding:}
\begin{itemize}
\item \reading{Reading assignments to be done before class on the scheduled day.}
\item \exam{Quizzes or exams set to take place on those days.}
\item \hw{Assignments due on these days.}
\item \noclass{No class on this day.}
\end{itemize}

\begin{calendar}{8/23/2021}{16} % Semester starts on 1/11/2010 and last for 16
                    % weeks, including finals week
\setlength{\calboxdepth}{.3in}
\setlength{\calwidth}{\textwidth}
\MTWFClass
%% schedule
\caltexton{1}{
\begin{itemize}
	\item First day. Syllabus and course material.
	\item Complex numbers.
	\item \reading{Review Chapter 1.}
	\item \reading{Chapter 3 Sections 1, 2.}
\end{itemize}
}

\caltextnext{
\begin{itemize}
\item Geometry of $\mathbb{C}$ and polar coordinates.
\item \reading{Chapter 3 Sections 3, 4.}
\end{itemize}
}

\caltextnext{
\begin{itemize}
	\item Polar coordinates and periodicity.
	\item \reading{Chapter 3 Sections 4, 5.}
\end{itemize}}

\caltextnext{
\begin{itemize}
	\item Intro to ODEs.
	\item \reading{Chapter 4 Section 1, 2.}
	\item \hw{Homework 0 due.}
	\item \hw{Discussion due: A mathematician's lament.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item General and particular solutions.
	\item \reading{Chapter 4 Sections 3.}
\end{itemize}  }

\caltextnext{
\begin{itemize}
	\item Separable ODEs.
	\item \reading{Chapter 4 Sections 4.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item \textcolor{purple}{Iodine clock experiment.}
		\item Changing variables and qualitative analysis.
		\item \reading{Chapter 4 Section 5, 6.}
\end{itemize}}

\caltextnext{
\begin{itemize}
	\item \exam{Quiz 1.}
	\item \hw{Homework 1 due.}
	\item \hw{Discussion due: The Mandelbrot and Julia Sets.}
	\item \hw{Review due: Homework 0.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item First order linear equations and integrating factor.
	\item \reading{Chapter 4 Section 7.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item Chemical kinetics.
	\item \reading{Chapter 4 Section 8.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item Cascading chemical reaction example.
	\item \hw{Discussion due: Is Mathematics Invented or Discovered?}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item  Second order ODEs and initial value problems.
	\item \reading{Chapter 4 Section 9.}
	\item \hw{Homework 2 due.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item Damped oscillation.
	\item \reading{Chapter 4 Section 9.}
	\item \hw{Review due: Homework 1.}
	\item \hw{Quiz Redo due: Quiz 1.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
    \item Driven and damped oscillation.
	\item \reading{Chapter 5 Section 1.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item \exam{Quiz 2}
	\item \hw{Discussion due: Richard Feynman.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item Boundary value problems and the Schr\"odinger equation.
	\item \reading{Chapter 5 Section 2.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item More on the Schr\"odinger equation.
	\item \reading{Chapter 5 Section 2.}
	\item \hw{Homework 3 due.}
	\item \hw{Review due: Homework 2.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item \exam{Exam 1.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item \exam{Exam 1.}
	\item \hw{Review due: Homework 3.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item Sequences and series.
	\item \reading{Chapter 6 Section 1, 2.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item Series and convergence.
    \item \textcolor{red}{Explicit Euler method.}
	\item \reading{Chapter 6 Section 2.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item Power series and radius of convergence.
	\item \reading{Chapter 7 Section 1.}
	\end{itemize}}

\caltextnext{
\begin{itemize}
	\item Continue.
	\item \reading{Chapter 7 Section 1.}
	\item \hw{Homework 4 due.}
	\item \hw{Discussion due: TBD.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Integration and differentiation with power series.
	\item \reading{Chapter 7 Section 2.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Taylor series.
	\item \reading{Chapter 7 Section 3.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Approximation with Taylor series and Morse potential.
	\item \reading{Chapter 7 Section 4.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item \exam{Quiz 3.}
	\item \hw{Homework 5 due.}
	\item \hw{Discussion due: TBD.}
	\item \hw{Review due: Homework 4.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Series solutions to ODEs.
	\item \reading{Chapter 7 Section 5.}
	\end{itemize}}
\caltextnext{
\begin{itemize}
	\item Continue.
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Special polynomials.
	\item \reading{Chapter 7 Section 6.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Quantum harmonic oscillator.
	\item \reading{Chapter 7 Section 7.}
	\item \hw{Homework 6 due.}
	\item \hw{Discussion due: TBD.}
	\item \hw{Review due: Homework 5.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Continue.
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Open.
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item \exam{Oral Exam 2.}
	\item \hw{Review due: Homework 6.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item \exam{Oral Exam 2.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Vectors and vector spaces.
	\item \reading{Chapter 8 Sections 1, 2.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Algebra of vector spaces.
	\item \reading{Chapter 8 Section 3, 4.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Inner and cross products.
	\item \reading{Chapter 8 Section 5.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Linear transformations and matrices.
	\item \reading{Chapter 9 Section 1.}
	\item \hw{Homework 7 due.}
	\item \hw{Discussion due: TBD.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Continue.
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Matrix algebra.
	\item \reading{Chapter 9 Section 2.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Systems of inhomogeneous linear equations.
	\item \reading{Chapter 9 Section 3, 4.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Systems of homogeneous equations, nullspace.
	\item \reading{Chapter 9 Section 3, 4.}
	\item \hw{Homework 8 due.}
	\item \hw{Discussion due: TBD.}
	\item \hw{Review due: Homework 7.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Linear independence, span, and bases.
	\item \reading{Chapter 9 Section 5.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Determinants, traces, and their properties.
	\item \reading{Chapter 9 Section 6.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Continue.
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item \exam{Quiz 4.}
	\item \hw{Homework 9 due.}
	\item \hw{Discussion due: TBD.}
	\item \hw{Review due: Homework 8.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Inverse and similar matrices.
	\item \reading{Chapter 9 Section 7.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Eigen-problem.
	\item \reading{Chapter 9 Section 8.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Diagonalization and Hermitian matrices.
	\item \reading{Chapter 9 Section 9.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Continue.
	\item \hw{Homework 10 due.}
	\item \hw{Discussion due: TBD.}
	\item \hw{Review due: Homework 9.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Groups and symmetries.
	\item \reading{Chapter 9 Section 10.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Continue.
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Continue.
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item \exam{Quiz 5.}
	\item \hw{Homework 11 due.}
	\item \hw{Discussion due: TBD.}
	\item \hw{Review due: Homework 10.}
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Project and review.
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item Project and review.
	\end{itemize} }

\caltextnext{
\begin{itemize}
	\item \exam{Exam 3.}
	\item \hw{Review due: Homework 11.}
	\end{itemize} }

	\caltextnext{
	\begin{itemize}
		\item \exam{Exam 3.}
		\end{itemize} }
%

% Holidays
\Holiday{9/6/2021}{\noclass{Labor Day}}
\Holiday{11/22/2021}{\noclass{Fall Break}}
\Holiday{11/23/2021}{\noclass{Fall Break}}
\Holiday{11/24/2021}{\noclass{Fall Break}}
\Holiday{11/26/2021}{\noclass{Fall Break}}

\end{calendar}
\end{center}
\end{document}