%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article} %report allows for chapters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{preamble}
\newcommand{\vhat}{\boldsymbol{\hat{v}}}

\begin{document}

\begin{center}
   \textsc{\large MATH 271, Homework 7, \emph{Solutions}}\\
   \textsc{Due November 1$^\textrm{st}$}
\end{center}
\vspace{.5cm}

\begin{problem}
Let $S$ be the set of general solutions $x(t)$ to the following homogeneous linear differential equation 
\[
x''+f(t)x'+g(t)x=0.
\]
Show that this set $S$ is a vector space over the complex numbers by doing the following. Let $x(t),y(t),z(t) \in S$ be solutions to the above equation and let $\alpha, \beta \in \C$ be complex scalars.
\begin{enumerate}[(a)]
    \item Write down the eight requirements for $S$ to be a vector space.  
    \item Identify the $\zerovec \in S$ and $1\in \C$. 
    \item Show that $\alpha x(t) + \beta y(t) \in S$. That is, show that a superposition of solutions is also a solution. \emph{Hint: We have shown this before.}
\end{enumerate}
\end{problem}
\begin{solution}
\begin{enumerate}[(a)]
    \item We can remember these requirements via the acronym CANI ADDU.  So we have for the vector addition properties
    \begin{itemize}
        \item Commutivity: If we have two solutions $x(t)$ and $y(t)$ in the set $S$, then we know 
        \[
        x(t)+y(t)=y(t)+x(t)
        \]
        is satisfied.
        \item Associativity: If we have three solutions $x(t),y(t),z(t)\in S$, then we know
        \[
        (x(t)+y(t))+z(t)=x(t) + (y(t)+z(t))
        \]
        is satisfied.
        \item Neutral Element: We have that there exists the zero function $0\in S$ such that
        \[
        0+x(t)=x(t).
        \]
        \item Inverses: Given an $x(t)\in S$, we have the function $-x(t)\in S$ such that
        \[
        x(t)+(-x(t))=0.
        \]
    \end{itemize}
    Then we have the scalar multiplication properties
    \begin{itemize}
        \item Associativity: If we have $\alpha,\beta \in \C$ and $x(t)\in S$ then we have
        \[
        \alpha (\beta x(t)) = (\alpha \beta)x(t)
        \]
        holds.
        \item Distribution: Given $\alpha,\beta \in \C$ and $x(t)\in S$ we have
        \[
        (\alpha +\beta)x(t) = \alpha x(t) + \beta x(t)
        \]
        holds.
        \item Distribution: Given $\alpha\in \C$ and $x(t),y(t)\in S$, we have
        \[
        \alpha(x(t)+y(t))=\alpha x(t) + \alpha y(t)
        \]
        holds.
        \item Unit element: We have $1\in \C$ satisfies that for any $x(t)\in S$ that
        \[
        1 x(t) = x(t).
        \]
    \end{itemize}
    \item Now, note that above we defined $\zerovec\in S$ to be the zero function 0. That is, the function that is 0 for every value of $t$. Note that $0$ is a solution to the equation since
    \[
    0'' + f(t)\cdot 0' + g(t) 0 = 0.
    \]
    Then, we have $1\in \C$ that satisfies the necessary property too. In this case, 1 is literally the unit element we care about.  
    
    \begin{remark} 
        Not all vector spaces will have such obvious neutral elements, $\zerovec$. Likewise, not all fields will have an obvious unit element $1$.  This is why we must be a bit careful at times.
    \end{remark}
    \item Now, the biggest requirement for a vector space is that linear combinations of vectors actually produce another vector. This is quite obvious in the plane $\R^2$ for example, but here, it is not necessarily obvious.  
    
    Now, we take $\alpha,\beta \in \C$ and $x(t),y(t)\in S$ and we consider the linear combination
    \[
    z(t) = \alpha x(t) + \beta y(t).
    \]
    We then wish to show that this linear combination (or superposition) is a solution as well. So we plug in $z(t)$ into our equation as follows
    \begin{align*}
        z''(t)+f(t)z'(t)+g(t)z(t)&= (\alpha x''(t) + \beta y''(t))+f(t)(\alpha x'(t) + \beta y'(t))+g(t)(\alpha x(t)+\beta y(t))\\
        &= \alpha \left[x''(t)+f(t)x'(t)+g(t)x(t)\right] + \beta \left[ y''(t)+f(t)y'(t)+g(t)y(t)\right]\\
        &=0,
    \end{align*}
    since we knew that $x(t)$ and $y(t)$ themselves are solutions. Thus, $z(t)$ is as well and now we have that $S$ is a vector space.
\end{enumerate}
\end{solution}

\newpage
\begin{problem}
Consider the following vectors in the real plane $\R^2$. We let
\[
\vecu = 1\xhat + 2\yhat \qquad \textrm{and} \qquad \vecv = -3\xhat+ 3\yhat.
\]
\begin{enumerate}[(a)]
    \item Draw both $\vecu$ and $\vecv$ in the plane and label the origin.
    \item Draw the vector $\vecw = \vecu+\vecv$ in the plane.
    \item Find the area of the parallelogram generated by $\vecu$ and $\vecv$.
\end{enumerate}
\end{problem}
\begin{solution}
\begin{enumerate}[(a)]
    \item See the plane below.
    \item Both (a) and (b) are in the plane here:
    \begin{center}
        \begin{tikzpicture}
        \draw[thin,gray!40] (-3,-3) grid (5,5);
        \draw[<->] (-3,0)--(5,0) node[right]{$x$};
        \draw[<->] (0,-3)--(0,5) node[above]{$y$};
        \draw[line width=2pt,blue,-stealth](0,0)--(1,2) node[anchor=west] at (1,2){$\vecu$};
        \draw[line width=2pt, red, -stealth](0,0)--(-3,3) node[anchor=north] at (-3,3){$\vecv$};
        \draw[line width=2pt, green, -stealth](0,0)--(-2,5) node[anchor=west] at (-2,5){$\vecu+\vecv$};
        \end{tikzpicture}
        \end{center}
        \item One could compute the area of the parallelogram generated by $\vecu$ and $\vecv$ in many ways. First, let us see what this looks like:
            \begin{center}
        \begin{tikzpicture}
        \draw[thin,gray!40] (-3,-3) grid (5,5);
        \draw[<->] (-3,0)--(5,0) node[right]{$x$};
        \draw[<->] (0,-3)--(0,5) node[above]{$y$};
        \draw[line width=2pt,blue,-stealth](0,0)--(1,2) node[anchor=west] at (1,2){$\vecu$};
        \draw[line width=2pt, red, -stealth](0,0)--(-3,3) node[anchor=north] at (-3,3){$\vecv$};
        \draw[line width=2pt, blue, -stealth](-3,3)--(-2,5) node[anchor=west] at (-2,5){};
        \draw[line width=2pt, red, -stealth](1,2)--(-2,5) node[anchor=west] at (-2,5){};
        \end{tikzpicture}
        \end{center}
        In order to compute this area, we can use the cross product by thinking of these vectors as being in $3$-dimensional space by
        \[
        \vecu = 1\xhat + 2\yhat + 0\zhat \qquad \textrm{and} \qquad \vecv = -3\xhat +3\yhat + 0\zhat.
        \]
        Then the cross product of these two vectors must only have a $z$-component since these two vectors lie in the $xy-$plane. Thus, we can compute
        \begin{align*}
            \vecu \times \vecv = (1\cdot 3 - (-3)\cdot2)\zhat = 9\zhat.
        \end{align*}
        Hence, the area is $\|\vecu\times \vecv\| = \|9\zhat \| = 9.$
\end{enumerate}
\end{solution}

\newpage
\begin{problem}~
\begin{enumerate}[(a)]
    \item We can reflect a vector in the plane by first reflecting basis vectors. Let $R\colon \R^2 \to \R^2$ be a function be defined by 
    \[
    R(\xhat) = -\xhat \qquad \textrm{and} \qquad R(\yhat)=\yhat.
    \]
    Let $\vecv = \alpha_1 \xhat + \alpha_2 \yhat$ and define
    \[
    R(\vecv) = \alpha_1 R(\xhat) + \alpha_2 R(\yhat).
    \]
    When this is the case, we call the function $T$ \underline{linear}.\\
    \noindent Show that $R$ reflects the vector $\vecu = 1\xhat + 2\yhat$ about the $y$-axis and draw a picture.
    \item We can rotate a vector in the plane by first rotating the basis vectors $\xhat$ and $\yhat$. Define a \underline{linear} function $T\colon \R^2 \to \R^2$ defined by
    \[
    T(\xhat)=\yhat \qquad \textrm{and} \qquad T(\yhat)=-\xhat.
    \]
    \noindent Show that $T$ rotates $\vecu$ by $\pi/2$ in the counterclockwise direction and draw a picture.
\end{enumerate}
\end{problem}
\begin{solution}~
\begin{enumerate}[(a)]
    \item So, we can take the vector $\vecu$ and then we have
    \[
    R(\vecu)=1R(\xhat)+2R(\yhat) = -1\xhat +2\yhat.
    \]
    So we can plot both $\vecu$ and $R(\vecu)$ in the plane:
        \begin{center}
        \begin{tikzpicture}
        \draw[thin,gray!40] (-3,-3) grid (3,3);
        \draw[<->] (-3,0)--(3,0) node[right]{$x$};
        \draw[<->] (0,-3)--(0,3) node[above]{$y$};
        \draw[line width=2pt,blue,-stealth](0,0)--(1,2) node[anchor=west] at (1,2){$\vecu$};
        \draw[line width=2pt, red, -stealth](0,0)--(-1,2) node[anchor=east] at (-1,2){$R(\vecu)$};
        \end{tikzpicture}
        \end{center}
        We can see that this is definitely the reflection of the vector $\vecu$ across the $y$-axis.
    \item We can now do this for the function $T$ to get
    \[
    T(\vecu) = 1T(\xhat)+2T(\yhat)=1\yhat -2\xhat.
    \]
    Then we can plot both $\vecu$ and $T(\vecu)$ in the plane:
        \begin{center}
        \begin{tikzpicture}
        \draw[thin,gray!40] (-3,-3) grid (3,3);
        \draw[<->] (-3,0)--(3,0) node[right]{$x$};
        \draw[<->] (0,-3)--(0,3) node[above]{$y$};
        \draw[line width=2pt,blue,-stealth](0,0)--(1,2) node[anchor=west] at (1,2){$\vecu$};
        \draw[line width=2pt, red, -stealth](0,0)--(-2,1) node[anchor=east] at (-2,1){$T(\vecu)$};
        \end{tikzpicture}
        \end{center}
        We can see that this is definitely the rotation  of the vector $\vecu$ an angle of $\pi/2$ in the counter-clockwise direction.
\end{enumerate}
\end{solution}

\newpage
\begin{problem}
Consider the following vectors in space $\R^3$
\[
\vecu = 1\xhat + 2\yhat + 3\zhat \qquad \textrm{and} \qquad \vecv = -2\xhat +1\yhat -2\zhat.
\]
\begin{enumerate}[(a)]
    \item Compute the dot product $\vecu\cdot \vecv$. 
    \item Compute the cross product $\vecu \times \vecv$.
    \item Compute the lengths $\|\vecu\|$ and $\|\vecv\|$ using the dot product.
    \item Compute the angle between vectors $\vecu$ and $\vecv$. 
    \item Compute the projection of $\vecu$ in the direction of $\vecv$. 
\end{enumerate}
\end{problem}
\begin{solution}~
\begin{enumerate}[(a)]
    \item We have that
    \begin{align*}
        \vecu \cdot \vecv &= 1\cdot (-2) + 2 \cdot 1 + 3 \cdot (-3)\\
        &= -6.
    \end{align*}
    \item Here, feel free to use a formula for a cross product instead of writing it all out.  We will find that
    \begin{align*}
        \vecu \times \vecv &= -7\xhat -4\yhat +5\zhat.
    \end{align*}
    \item We compute the lengths using the dot product by
    \[
        \|\vecu\| = \sqrt{\vecu \cdot \vecu} = \sqrt{1^2+2^2+3^2} = \sqrt{14}.
    \]
    Likewise
    \[
        \|\vecv\| = \sqrt{\vecv \cdot \vecv} = \sqrt{(-2)^2 + 1^2 (-2)^2} = \sqrt{9} = 3.
    \]
    \item We can find the angle $\theta$ between $\vecu$ and $\vecv$ with the information we already have.  Indeed, recall the formula
    \[
        \vecu \cdot \vecv = \|\vecu \| \|\vecv\| \cos \theta.
    \]
    In the previous parts, we have computed $\vecu \cdot \vecv = -6$, $\|\vecu\|=\sqrt{14}$, and $\|\vecv\|=3$ so we now just have the following
    \[
        -6 = 3\sqrt{14} \cos \theta.
    \]
    Thus,
    \[
        \theta = \arccos\left(\frac{-6}{3\sqrt{14}}\right) \approx 2.1347 ~\textrm{radians}.
    \]
    One could also use the fact that $\|\vecu \times \vecv\| = \|\vecu\| \|\vecv\| \sin \theta$ since we had already compute the cross product earlier. 
    \item The projection of $\vecu$ in the direction of $\vecv$ is simply asking for how much of the vector $\vecu$ is in the direction of $\vecv$.  One can arrive at this purely through trigonometry, but we have the dot product at our disposal.  The normalized vector $\vhat$ points in the direction of $\vecv$ with length 1 and
    \[
        \vhat = \frac{1}{\|\vecv\|} \vecv = \frac{1}{3} \vecv.
    \]
    Then, the projection can be computed by
    \[
        \vecu \cdot \vhat = \frac{1}{3} \vecu \cdot \vecv = -2.
    \]
    One should attempt to recover this notion by doing some trigonometry. 
\end{enumerate}
\end{solution}
\end{document}