%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article} %report allows for chapters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{preamble}
\newcommand{\vecx}{\boldsymbol{\vec{x}}}
\newcommand{\vecy}{\boldsymbol{\vec{y}}}

\begin{document}

\begin{center}
   \textsc{\large MATH 271, Homework 10}\\
   \textsc{Due November 20$^\textrm{nd}$}
\end{center}
\vspace{.5cm}


\begin{problem}
For each description below, construct a single $3\times 3$ transformation matrix.
\begin{enumerate}[(a)]
    \item A rotation by $\frac{\pi}{2}$ in the $yz$-plane.
    \item A rotation by $\frac{\pi}{2}$ in the $xy$-plane, an interchange of the $x$ and $y$ coordinates.
    \item A matrix which undoes everything in part (b).
\end{enumerate}
\end{problem}


\begin{problem}
Consider the system of linear equations
\begin{align*}
    2x-y-2z&=4\\
    -x + 3y +4z &= -2\\
    2x+y-2z&=8.
\end{align*}
\begin{enumerate}[(a)]
    \item Write this system of equations as a matrix/vector equation
    \[
    [A]\vecx=\vecy.
    \]
    \item Solve the system of equations by finding the inverse of the matrix $[A]$.
\end{enumerate}
\end{problem}


\begin{problem}
Consider the matrix
\[
[A]=\begin{pmatrix} 1 & 3 & 0 \\ 3 & 1 & 0 \\
 0 & 4 & 1 \end{pmatrix}
\]
\begin{enumerate}[(a)]
    \item Compute the eigenvalues and eigenvectors for the matrix $[A]$.
    \item Show that $\det([A])=\lambda_1\lambda_2\lambda_3$ and $\tr([A])=\lambda_1 + \lambda_2+\lambda_3$ where $\lambda_1$, $\lambda_2$, and $\lambda_3$ are the eigenvalues you found in (a).
    \item Argue why the eigenvalues of $[A]^{-1}$ must be $\frac{1}{\lambda_1}$, $\frac{1}{\lambda_2}$, and $\frac{1}{\lambda_3}$.
\end{enumerate}
\end{problem}

\begin{problem}
Consider the matrix
\[
[A] = \begin{pmatrix} 1 & 1 \\ 0 & 2 \end{pmatrix}.
\]
\begin{enumerate}[(a)]
    \item Find the eigenvalues and eigenvectors for this matrix.
    \item Construct the matrix $[P]$ such that
    \[
    [\Lambda] = [P]^{-1}[A][P]
    \]
    from the eigenvectors you found. 
    \item Find $[P]^{-1}$ and compute
    \[
    [\Lambda] = [P]^{-1}[A][P].
    \]
    Is this $[\Lambda]$ diagonal?
\end{enumerate} 
\end{problem}

\begin{problem}
Let $A \colon \C^n \to \C^n$ be hermitian.  We prove the two theorems in the text.
\begin{enumerate}[(a)]
    \item Show that all eigenvalues of $A$ are real.
    \item Show that eigenvectors corresponding to different eigenvalues are orthogonal with the hermitian inner product.
\end{enumerate}
\end{problem}




\end{document}