%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article} %report allows for chapters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{preamble}

\begin{document}

\begin{center}
   \textsc{\large MATH 271, Homework 1}\\
   \textsc{Due September 3$^\textrm{rd}$}
\end{center}
\vspace{.5cm}

\begin{problem}
    Look up how to do \textbf{\emph{integration by parts}}. Use this technique to compute the integral
    \[
        \int t e^{3t}dt.
    \]
    
\end{problem}

\begin{problem}
Convert the following numbers in Cartesian coordinates to polar coordinates and compute all pairwise products after converting.
\begin{enumerate}[(a)]
    \item $z_1=\frac{1}{2}-\frac{1}{2}i$;
    \item $z_2=-1+3i$;
    \item $z_3=-2-3i$.
\end{enumerate}
\end{problem}

\begin{problem}
Find the square roots of $-i$ using a geometrical argument. \emph{Hint: think of polar coordinates and rotation!}
\end{problem}

\begin{problem}
Draw the unit circle in the complex plane. Plot the complex numbers $z_1$ and $z_2$ given above and find their inverses and conjugates. Explain what taking the inverse and conjugate does geometrically.
\end{problem}

\begin{problem}
    What is an (ordinary) differential equation? Explain what it means to be a general and particular solution to a differential equation.
\end{problem}

\begin{problem}
    Look up an ordinary differential equation in chemistry that interests you.  Write it down, and explain what it attempts to model. Why does it interest you?
\end{problem}

\begin{problem}
Objects near Earth fall due to the force of gravity.  The acceleration of an object due to gravity (regardles of mass) is then
\[
y''=g,
\]
where $y(t)$ represents the height above the ground at time $t$ and $g\approx -9.8\frac{m}{s^2}$ is constant.  
\begin{enumerate}[(a)]
    \item Find the general solution to the equation.
    \item Given the initial data $y(0)=0$ and $y'(0)=1$, find the particular solution.
    \item At what time $t>0$ does the object first contact the ground?
    \item Plot your solution only over the range of time that makes physical sense. 
\end{enumerate}
\end{problem}

\end{document}