%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article} %report allows for chapters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{preamble}

\begin{document}

\begin{center}
   \textsc{\large MATH 271, Homework 0}\\
   \textsc{Due August 28$^\textrm{th}$}
\end{center}

\begin{center}
    \textbf{Tell me a little bit about yourself!}
\end{center}

\noindent For the following five questions, write a message in our teams chat in the ``About you" channel.  Put your answer to all five questions in one single message please! This will make it easier to read.

\begin{question}
    Where did you grow up? Why did you choose Colorado State University?
\end{question}

\begin{question}
    What is something interesting about yourself?  What is a favorite hobby of yours?
\end{question}

\begin{question}
    Why are you interested in chemistry? Do you have plans after you earn your degree?
\end{question}

\begin{question}
    \emph{Honest answers appreciated here.}  Do you enjoy mathematics? Do you think it's easy or hard? What do you hope to gain from this course?
\end{question}

\begin{question}
    Why did you choose Math 271 and 272 as opposed to the physical sciences calculus sequence?
\end{question}


\begin{center}
    \textbf{Now for some mathematics.}
\end{center}

\begin{problem}
    Compute the following:
\begin{enumerate}[(a)]
    \item $\displaystyle{\frac{d}{dx}(2x^7-3x^4+7)}$;
    \item $\displaystyle{\frac{d}{dt}\left(e^{at}\sin(bt)\right)}$;
    \item $\displaystyle{\frac{d}{ds}\left(\tan\left( e^{s^2}\right)\right)}$.
\end{enumerate}
\end{problem}

\begin{problem}
    Compute the following:
\begin{enumerate}[(a)]
    \item $\displaystyle{\int 2x^7-3x^4+7dx}$;
    \item $\displaystyle{\int_{-1}^1 \cos(t) dt}$;
    \item $\displaystyle{\int_0^1 e^{xy}dy}$;
\end{enumerate}
\end{problem}

\begin{problem}
    Find the point(s) of intersection of the parabola $f(x)=2x^2+2x+2$ and the line $g(x)=4x+4$ using algebra. Plot this on Desmos to confirm your answer and include a picture of the two graphs.
\end{problem}

\begin{problem}
    Now take the same parabola $f(x)=2x^2+2x+2$ and the line $h(x)=4x-4$. Plot this on Desmos and include a picture to show that the graphs of these two functions do not intersect in the plane. Explain why this must be true.
    
    We can find ``complex intersections" by doing the same algebra as the previous problem. Find these complex interesctions. (\emph{Hint: set up an equation whose roots would give you the intersections of these two curves.)}
\end{problem}


\end{document}