%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article} %report allows for chapters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{preamble}

\begin{document}

\begin{center}
   \textsc{\large MATH 271, Homework 2}\\
   \textsc{Due September 13$^\textrm{th}$}
\end{center}
\vspace{.5cm}

\begin{problem}
    Solve the following autonomous equation.
    \[
    x'=-x^2.
    \]
    Can $x(0)=0$ be an initial condition?
\end{problem}

\begin{problem}
Objects near Earth fall due to gravity.  The acceleration of an object due to gravity is then
\[
x''=g,
\]
where $x$ represents the distance above the ground and $g\approx -9.8\frac{m}{s^2}$.  
\begin{enumerate}[(a)]
    \item Find the general solution to the equation.
    \item Given the initial data $x(0)=0$ and $x'(0)=1$, find the particular solution.
    \item Plot your solution over a meaningful range of time.
    \item When is the object touching the ground?
\end{enumerate}

\end{problem}

\begin{problem}
Consider the following differential equation.
\[
x' = x\cos(t).
\]
\begin{enumerate}[(a)]
    \item What is the order of this equation?
    \item Find the general solution to this equation.
    \item Given the initial data $x(0)=1$, find the particular solution.
    \item Plot this function and explain in words what the solution represents if $x(t)$ is position.
\end{enumerate}
\end{problem}

\begin{problem}
Consider the differential equation
\[
x'=\frac{x+t}{t}.
\]
\begin{enumerate}[(a)]
    \item Let $f(x,t)=\frac{x+t}{t}$. Show that $f(x,t)=f(\lambda x, \lambda t)$.
    \item Given (a) holds, use the change of variables $u=\frac{x}{t}$ to rewrite the differential equation as a separable equation in terms of $u$.
    \item Find the general solution to the equation and write your solution in terms of the original variables $t$ and $x$.
\end{enumerate}
\end{problem}

\begin{problem}
Find the general solution to the following equation.
\[
tx'+2x=\frac{\sin(t)}{t}.
\]
Show that your solution is correct. (\emph{Hint: can you use an integrating factor?})
\end{problem}


\end{document}