%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article} %report allows for chapters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{preamble}

\begin{document}

\begin{center}
   \textsc{\large MATH 271, Worksheet 2}\\
   \textsc{Ordinary Differential Equations}
\end{center}
\vspace{.5cm}

\begin{center}
    Problems 1-6 are related.
\end{center}

\begin{problem}
    (Newton's law of cooling) Write down a differential equation that models the following scenario:\\
    
    \noindent\emph{The temperature of a substance in an ambient environment changes over time proportionally to the difference of the substance from the ambient environment.}\\
    
    \noindent Let $T(t)$ be the temperature of the substance, $T_a$ be the ambient temperature, and $k$ be the constant of proportionality.
\end{problem}

\begin{problem}
    With the equation found above, find a general solution.
\end{problem}

\begin{problem}
    With the parameter values $T_a=100$, $k=1$, and initial data $T(0)=50$, find the particular solution.
\end{problem}

\begin{problem}
    Plot the particular solution to the problem and interpret what this means for the temperature over time.
\end{problem}

\begin{problem}
    What happens instead if the initial temperature is equal to the ambient temperature? Does your solution reflect this? Does this make physical sense?
\end{problem}

\begin{problem}
    Let $\delta = (T_a-T)$. Show that the equation you found in Problem 1 reduces to
    \[
    \delta' = -k\delta.
    \]
    What is this equation describing physically?
\end{problem}

\begin{center}
    Problems 7-8 are related.
\end{center}
\begin{problem}
    Show that $x=c_1\sin(t)+c_2\cos(t)$ is a general solution to the equation
    \[
        x''+x=0.
    \]
\end{problem}

\begin{problem}
    Find the particular solution if $x(0)=1$ and $x'(0)=2$. Plot your solution in the $t,x$-plane and in the $x,x'$-plane. 
\end{problem}

\begin{problem}
    Consider the differential equation
    \[
    x'=\frac{x^2+tx+t^2}{tx}.
    \]
    \begin{enumerate}[(a)]
        \item Let $f(x,t)=x^2+tx+t^2$.  Show that $f(\lambda x, \lambda t)=f(x,t)$.
        \item Use the substitution $u=\frac{x}{t}$ in order to make the original equation separable.
        \item Find the general solution to this separable equation in terms of $u$ and $t$. You may use Wolfram Alpha to compute the necessary integral.
        \item Find the solution to the original equation using the substitution $u=\frac{x}{t}$ and your solution from (c).
    \end{enumerate}
\end{problem}


\end{document}
