\documentclass{article}
\usepackage{termcal}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{xcolor}
% Few useful commands (our classes always meet either on Monday and Wednesday 
% or on Tuesday and Thursday)

\newcommand{\MTWFClass}{%
\calday[Monday]{\classday} % Monday
\calday[Tuesday]{\classday} % Tuesday
\calday[Wednesday]{\classday} % Wednesday
\skipday % Thursday (no class)
\calday[Friday]{\classday} % Friday 
\skipday\skipday % weekend (no class)
}

\newcommand{\TRClass}{%
\skipday % Monday (no class)
\calday[Tuesday]{\classday} % Tuesday
\skipday % Wednesday (no class)
\calday[Thursday]{\classday} % Thursday
\skipday % Friday 
\skipday\skipday % weekend (no class)
}

\newcommand{\Holiday}[2]{%
\options{#1}{\noclassday}
\caltext{#1}{#2}
}

\begin{document}
\begin{center}
   \textsc{\large MATH 271, Calendar}\\
   \textsc{Fall 2020}
\vspace{.5cm}

\begin{calendar}{8/24/2020}{16} % Semester starts on 1/11/2010 and last for 16
                    % weeks, including finals week
\setlength{\calboxdepth}{.3in}
\MTWFClass
%% schedule
\caltexton{1}{First day. Syllabus and course material. \textcolor{green}{Review Chapter 1.}}
\caltextnext{Complex numbers. \textcolor{green}{Chapter 3 Sections 1, 2.}}
\caltextnext{Geometry of $\mathbb{C}$ and polar coordinates. \textcolor{green}{Chapter 3 Sections 3, 4.}}
\caltextnext{\textbf{\textcolor{red}{Homework 0 due.}} Polar coordinates and periodicity. \textcolor{green}{Chapter 3 Sections 4, 5.}}
\caltextnext{Intro to ODEs. \textcolor{green}{Chapter 4 Section 1, 2.}}
\caltextnext{General and particular solutions. Separable ODEs. \textcolor{green}{Chapter 4 Sections 3, 4.}}
\caltextnext{Changing variables and qualitative analysis. \textcolor{green}{Chapter 4 Section 5, 6.}}
\caltextnext{\textbf{\textcolor{red}{Quiz 1. Homework 1 due.}}}
\caltextnext{First order linear equations and integrating factor. \textcolor{green}{Chapter 4 Section 7.}}
\caltextnext{Chemical kinetics. \textcolor{green}{Chapter 4 Section 8.}}
\caltextnext{\textbf{\textcolor{red}{Homework 2 due.}} Second order ODEs and initial value problems. \textcolor{green}{Chapter 4 Section 9}}
\caltextnext{Cont.\textcolor{green}{Chapter 4 Section 9.}}
\caltextnext{Damped and driven oscillation. \textcolor{green}{Chapter 4 Section 9.}}
\caltextnext{Boundary value problems. \textcolor{green}{Chapter 5 Section 1.}}
\caltextnext{\textbf{\textcolor{red}{Quiz 2}}}
\caltextnext{\textbf{\textcolor{red}{Homework 3 due.}} Understanding the Schr\"odinger equation. \textcolor{green}{Chapter 5 Section 2.}}
\caltextnext{More on the Schr\"odinger equation. \textcolor{green}{Chapter 5 Section 2.}}
\caltextnext{\textbf{\textcolor{red}{Oral Exam 1}}}
\caltextnext{\textbf{\textcolor{red}{Oral Exam 1}}}
\caltextnext{Sequences and series. \textcolor{green}{Chapter 6 Section 1, 2.}}
\caltextnext{Series and convergence. \textcolor{green}{Chapter 6 Section 2.}}
\caltextnext{Power series and radius of convergence. \textcolor{green}{Chapter 7 Section 1.}}
\caltextnext{Cont. \textcolor{green}{Chapter 7 Section 1.} \textbf{\textcolor{red}{Homework 4 due.}}}
\caltextnext{Integration and differentiation with power series. \textcolor{green}{Chapter 7 Section 2.}}
\caltextnext{Taylor series. \textcolor{green}{Chapter 7 Section 3.}}
\caltextnext{Approximation with Taylor series and Morse potential. \textcolor{green}{Chapter 7 Section 4.}}
\caltextnext{\textbf{\textcolor{red}{Quiz 3. Homework 5 due.}}}
\caltextnext{Series solutions to ODEs. \textcolor{green}{Chapter 7 Section 5.}}
\caltextnext{Cont.}
\caltextnext{Special polynomials. \textcolor{green}{Chapter 7 Section 6.}}
\caltextnext{Quantum harmonic oscillator.}
\caltextnext{\textbf{\textcolor{red}{ Homework 6 due.}} Cont.}
\caltextnext{Open}
\caltextnext{\textbf{\textcolor{red}{Oral Exam 2}}}
\caltextnext{\textbf{\textcolor{red}{Oral Exam 2}}}
\caltextnext{\textcolor{purple}{SNOW DAY.} Vectors and vector spaces. \textcolor{green}{Chapter 8 Sections 1, 2.}}
\caltextnext{Algebra of vector spaces. \textcolor{green}{Chapter 8 Section 3, 4.}}
\caltextnext{Inner and cross products. \textcolor{green}{Chapter 8 Section 5.}}
\caltextnext{Linear transformations and matrices. \textcolor{green}{Chapter 9 Section 1.}} 
\caltextnext{Cont. \textbf{\textcolor{red}{Homework 7 due.}}}
\caltextnext{Matrix algebra. \textcolor{green}{Chapter 9 Section 2.}}
\caltextnext{Systems of inhomogeneous linear equations. \textcolor{green}{Chapter 9 Section 3, 4.}}
\caltextnext{\textbf{\textcolor{red}{Homework 8 due.}} Systems of homogeneous equations, nullspace.  \textcolor{green}{Chapter 9 Section 3, 4.}}
\caltextnext{Linear independence, span, and bases. \textcolor{green}{Chapter 9 Section 5.}}
\caltextnext{Determinants, traces, and their properties. \textcolor{green}{Chapter 9 Section 6.}}
\caltextnext{Cont.}
\caltextnext{\textbf{\textcolor{red}{Quiz 4. Homework 9 due.}}}
\caltextnext{Inverse and similar matrices. \textcolor{green}{Chapter 9 Section 7.}}
\caltextnext{Eigen-problem. \textcolor{green}{Chapter 9 Section 8.}}
\caltextnext{Diagonalization and Hermitian matrices. \textcolor{green}{Chapter 9 Section 9.}}
\caltextnext{\textbf{\textcolor{red}{Homework 10 due.}} Cont.}
\caltextnext{Groups and symmetries. \textcolor{green}{Chapter 9 Section 10.}}
\caltextnext{Cont.}
\caltextnext{Cont.}
\caltextnext{\textbf{\textcolor{red}{Quiz 5. Homework 11 due.}}}
\caltextnext{Project and review.}
\caltextnext{Project and review.}
\caltextnext{\textbf{\textcolor{red}{Oral Exam 3}}}
\caltextnext{\textbf{\textcolor{red}{Oral Exam 3}}}
%

% Holidays
\Holiday{9/7/2020}{\textcolor{blue}{Labor Day}}
\Holiday{11/23/2020}{\textcolor{blue}{Fall Break}}
\Holiday{11/24/2020}{\textcolor{blue}{Fall Break}}
\Holiday{11/25/2020}{\textcolor{blue}{Fall Break}}
\Holiday{11/27/2020}{\textcolor{blue}{Fall Break}}

\end{calendar}
\end{center}
\end{document}