\documentclass[12pt]{amsbook}
\usepackage{preamble}


\begin{document}
\pagenumbering{gobble}       % This kills the page numbering

\begin{center}
   \textsc{\large MATH 271, Exam 1}\\
   \textsc{Take Home Portion}\\
   \textsc{Due September 25$^\textrm{th}$ at the start of class}
\end{center}
\vspace{1cm}

\noindent\textbf{Name} \; \underline{\hspace{8cm}}

\vspace{1cm}

\noindent\textbf{Instructions} \; You are allowed a textbook, homework, notes, worksheets, material on our Canvas page, but no other online resources (including calculators or WolframAlpha). for this portion of the exam.  \textbf{Do not discuss any problem any other person.} All of your solutions should be easily identifiable and supporting work must be shown.  Ambiguous or illegible answers will not be counted as correct. \textbf{Print out this sheet and staple your solutions to it. Use a new page for each problem.}


\vspace{1cm}

\begin{center}\textbf{Problem 1} \; \underline{\hspace{1cm}}/15 \qquad \qquad \textbf{Problem 2} \; \underline{\hspace{1cm}}/10\end{center}

\vspace{1cm}

\hrule

\vspace*{1cm}
\noindent\emph{Note, these problems span two pages.}

\begin{problem}
Consider the endothermic breakdown of a molecule $x$ given by
\[
x \xrightarrow{kt^2} \textrm{Products}
\]
where we let $x(t)$ denote the concentration of reactants. Since the reaction is endothermic, if we also heat up the solution over time, we get a factor of $t^2$ as well since the reaction occurs more readily in higher temperatures. The concentration decreases over time based on differential equation
\[
x'=-kt^2(x-x_e).
\]
where $x_e$ is a constant that denotes the equilibrium concentration. 
\begin{enumerate}[(a)]
    \item Write an equivalent equation with the change of variables $\delta=x-x_e$.
    \item Find the general solution to this new equation. 
    \item What is the general solution in terms of the original variables $x$?
    \item Given the initial amount of $x$ is $x(0)=1$, the equilibrium concentration is $x_e=1/2$, and $k=1$, find the particular solution for $x(t)$.
    \item Does this reaction ever reach the equilibrium state?
\end{enumerate}
\end{problem}

\newpage
\begin{problem}
Consider the inhomogeneous linear differential equation
\[
x''+\omega^2x=\cos(\omega t),
\]
where $x(t)$ is a function of $t$. This is an example of resonance.
\begin{enumerate}[(a)]
    \item Find the general homogeneous solution $x_h(t)$.
    \item Show that your solution solves the homogeneous equation
    \[
    x_h''+\omega^2x_h=0.
    \]
    \item For the particular integral, $x_p(t)$, take the ansatz
    \[
    x_p(t)=C_1t\cos(\omega t) + C_2 t\sin(\omega t)
    \]
    and find the undetermined coefficients $C_1$ and $C_2$.
    \item Show that $x=x_h+x_p$ solves the inhomogeneous equation
    \[
    x''+\omega^2 x = \cos(\omega t).
    \]
\end{enumerate}
\end{problem}








\end{document}  