%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article} %report allows for chapters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{preamble}

\begin{document}

\begin{center}
   \textsc{\large MATH 271, Homework 8, \emph{Solutions}}\\
\end{center}
\vspace{.5cm}





\newpage
\begin{problem}
Consider the linear transformation $J \colon R^2 \to \R^2$ defined by
\[
J(\xhat) = \yhat \qquad \textrm{and} \qquad J(\yhat) = -\xhat.
\]
This linear transformation is fundamental in understanding how we can reconstruct complex numbers using matrices.
\begin{enumerate}[(a)]
    \item Show that $J^2 = J\circ J= -1$.
    \item Determine a matrix representation for $J$ and denote it by $[J]$.
    \item Recall that we can represent a complex number as $z=x + iy$ and that we can represent $z$ as a vector in $\R^2$ as $\vec{\boldsymbol{\zeta}} = x\xhat + y\yhat$.  Show that $J \vec{\boldsymbol{\zeta}}$ corresponds to $iz$.
    \item We can completely reconstruct a representation of $\C$ by using a matrix representation.  In particular, we can take
    \[
        [z] = x [I] + y [J].
    \]
    Show that we recover the complex addition and multiplication using this representation.
    \item We can represent a unit complex number as $z=e^{i\theta}$.  Show that the representation described before leads to
    \[
        [z] = \begin{pmatrix} \cos(\theta) & -\sin(\theta) \\ \sin(\theta) & \cos(\theta) \end{pmatrix}.
    \]
\end{enumerate}
\end{problem}
\begin{solution}
    \begin{enumerate}[(a)]
        \item Let $\vecv = v_1 \xhat + v_2 \yhat$ be some arbitrary vector in $\R^2$.  Then,
        \begin{align*}
            J^2(\vecv) = J(J(\vecv)) &= J(J(v_1 \xhat + v_2 \yhat))\\
            &= J(v_1 J(\xhat) + v_2 J(\yhat))\\
            &= J(v_1 \yhat - v_2 \xhat)\\
            &= v_1 J(\yhat) - v_2J(\xhat)\\
            &= -v_1 \xhat - v_2 \yhat\\
            &= -\vecv.
        \end{align*}
        So, yes, $J^2$ acts like scaling by -1.
        \item We determine a matrix for $J$ by using the definition of $J$ on $\xhat$ and $\yhat$. In particular,
        \[
            [J] = \begin{pmatrix} \vert & \vert \\ J(\xhat) & J(\yhat) \\ \vert & \vert \end{pmatrix} = \begin{pmatrix} 0 & -1 \\ 1 & 0 \end{pmatrix}.
        \]
        One can check here that $[J]^2=-[I]$, where $[I]$ is the identity matrix. This confirms that $[J]$ satisfies the relationship we saw in $(a)$.
        \item In the complex plane, we let $z=x+iy$ and we can note that
        \[
        iz = -y + ix.
        \]
        Now, we can think of $z$ as a vector in $\R^2$ by noticing that the vector $\vec{\boldsymbol{\zeta}} = x\xhat + y\yhat$ corresponds to the same exact point geometrically.  Then, if we apply $J$ we have
        \[
        J\vec{\boldsymbol{\zeta}} = -y\xhat + x \yhat,
        \]
        which is exactly how $z$ was transformed when we multiplied by $i$. Keep in mind that $i$ rotates a complex number $z$ by $\pi/2$ in the counterclockwise direction and $J$ does the same to vectors $\vec{\boldsymbol{\zeta}}$. To see this most fully, consider drawing a picture of both transformations.
        \item In the complex plane, we can take two complex numbers $z_1 = x_1 + iy_1$ and $z_2 = x_2 + i y_2$. Then we have
        \[
        z_1 + z_2 = (x_1 + x_2) + i(y_1 + y_2) \qquad \textrm{and} \qquad z_1 z_2 = (x_1x_2 - y_1 y_2) + i(x_1 y_2 + x_2 y_1).
        \]
        Notice that addition is componentwise and keep track of this result from the multiplication.

        Now, we can consider two matrices $[z_1] = x_1 [I] + y_1 [J]$ and $[z_2] = x_2 [I] + y_2 [J]$ and see what we get through addition and multiplication. We have
        \[
        [z_1] + [z_2] = (x_1 + x_2)[I] + (y_1 + y_2)[J].
        \]
        This is due to how matrices add componentwise and we can see that this corresponds to the addition in $\C$. Next, we have
        \begin{align*}
        [z_1][z_2] &= (x_1 [I] + y_1 [J])(x_2 [I] + y_2[J]) \\
        &= x_1 x_2 [I]^2 + y_1 x_2 [J][I] + x_1 y_2 [I][J] + y_1 y_2 [J]^2\\
        &= x_1 x_2 [I] + y_1 x_2 [J] + x_1 y_2 [J] - y_1 y_2 [I]\\
        &= (x_1 x_2 - y_1 y_2)[I] + (x_1 y_2 + x_2 y_1)[J].
        \end{align*}
        Note that I use the facts $[J][I]=[I][J]=[J]$, $[I]^2=[I]$, and from (a) we know $[J]^2=-[I]$.  If we take a look at the end result, we can see that this is the same multiplication result as $z_1 z_2$ in $\C$.

        \item Using our knowledge from the previous problem, and Euler's formula, we know that we can take
        \[
        [e^{i\theta}] = \cos (\theta) [I] + \sin(\theta)[J].
        \]
        Writing out the matrices explicitly yields
        \[
        [e^{i\theta}] = \begin{pmatrix} \cos(\theta) & 0 \\ 0 & \cos(\theta) \end{pmatrix} + \begin{pmatrix} 0 & - \sin(\theta) \\ \sin(\theta) & 0 \end{pmatrix} = \begin{pmatrix} \cos(\theta) & -\sin(\theta) \\ \sin(\theta) & \cos(\theta) \end{pmatrix},
        \]
        as intended.
    \end{enumerate}
\begin{remark}
    If one goes to look up a rotation matrix for $\R^2$, you will find the matrix you found in (e).  So, this goes to show that complex arithmetic captures rotations nicely through Euler's formula. Moreover, the matrix representation for a complex number is faithful in describing all that we need.
\end{remark}
\end{solution}

\newpage
\begin{problem}
Write down the matrix for the following linear transformation $T\colon \R^3 \to \R^3$:
\[
T\begin{pmatrix} x\\ y\\ z \end{pmatrix}
= \begin{pmatrix} x+y+z\\ 2x\\ 3y + z \end{pmatrix}.
\]
\end{problem}
\begin{solution}
We need that
\begin{align*}
[T] \begin{pmatrix} x \\ y \\ z \end{pmatrix} &= \begin{pmatrix} x+y+z\\ 2x\\ 3y + z \end{pmatrix}
\end{align*}
via matrix multiplication. Since the input vector is a 3-dimensional vector, and the output vector is $3$-dimensional, we must have that $[T]$ is a $3\times 3$-matrix. Hence,
\[
[T] = \begin{pmatrix} t_{11} & t_{12} & t_{13} \\ t_{21} & t_{22} & t_{23} \\ t_{31} & t_{32} & t_{33} \end{pmatrix}.
\]
Then we have
\begin{align*}
\begin{pmatrix} t_{11} & t_{12} & t_{13} \\ t_{21} & t_{22} & t_{23} \\ t_{31} & t_{32} & t_{33} \end{pmatrix} \begin{pmatrix} x \\ y \\ z \end{pmatrix} &=  \begin{pmatrix} t_{11}x + t_{12}y + t_{13}z \\ t_{21}x +  t_{22}y + t_{23}z \\ t_{31}x + t_{32}y + t_{33}z \end{pmatrix} = \begin{pmatrix} x + y + z \\ 2x \\ 3y+z \end{pmatrix}.
\end{align*}
If we match the coefficients on the $x$, $y$, and $z$, we find that
\[
[T] = \begin{pmatrix} 1 & 1 & 1 \\ 2 & 0 & 0 \\ 0 & 3 & 1 \end{pmatrix}.
\]
\end{solution}


\newpage
\begin{problem}
    Take the following matrices:
    \[
        [A] = \begin{pmatrix} 4 & 3 & 10 & 2 \\ 1 & 1 & 0 & 9 \end{pmatrix},\quad [B] = \begin{pmatrix} 8 & 5 & 8 \\ 10 & 9 & 2 \\ 4 & 6 &3 \end{pmatrix}, \quad [C] = \begin{pmatrix} 0 & 0 & 9 \\ 7 & 9 & 9 \\ 1 & 9 & 9 \\ 3 & 3 & 1\end{pmatrix}
    \]
    \begin{enumerate}[(a)]
        \item Compute either $[A][C]$ or $[C][A]$ and state which multiplication is not possible.
        \item Compute either $[B][C]$ or $[C][B]$ and state which multiplication is not possible.
        \item Can you add any of these matrices?
        \item Describe each matrix as linear transformation $T\colon \R^m \to \R^n$. What is $m$ and $n$ for each? How does this relate to the number of rows and columns?
    \end{enumerate}
\end{problem}
\begin{solution}~
\begin{enumerate}[(a)]
    \item The matrix $[A]$ is a $2\times 4$ matrix and matrix $[C]$ is a $4 \times 3$ matrix.  So we can compute $[A][C]$ but not $[C][A]$.  Given that, we also expect the output to be a $2 \times 3$ matrix. So, we have
    \[
    [A][C] = \begin{pmatrix} 4 & 3 & 10 & 2 \\ 1 & 1 & 0 & 9 \end{pmatrix}\begin{pmatrix} 0 & 0 & 9 \\ 7 & 9 & 9 \\ 1 & 9 & 9 \\ 3 & 3 & 1\end{pmatrix} = \begin{pmatrix} 37 & 123 & 155 \\ 34 & 36 & 27 \end{pmatrix}.
    \]

    \item $[B]$ is a $3 \times 3$ matrix so we can take $[C][B]$ but not $[B][C]$.  We get
    \[
        [C][B] = \begin{pmatrix} 0 & 0 & 9 \\ 7 & 9 & 9 \\ 1 & 9 & 9 \\ 3 & 3 & 1\end{pmatrix} \begin{pmatrix} 8 & 5 & 8 \\ 10 & 9 & 2 \\ 4 & 6 &3 \end{pmatrix} = \begin{pmatrix} 36 & 54 & 27 \\ 182 & 170 & 101 \\ 134 & 140 & 53 \\ 58 & 48 & 33 \end{pmatrix}.
    \]
    \item We can always add a matrix to itself, so, for example $[A]+[A]$, $[B]+[B]$, and $[C]+[C]$ make sense. However, since the dimensions of $[A]$, $[B]$, and $[C]$ all differ, we cannot add in any other way.
    \item The number of columns of a matrix denotes the input dimension $m$, and the number of rows denotes the output dimension $n$.  So
    \[
        A \colon \R^4 \to \R^2, \quad B \colon \R^3 \to \R^3, \quad C \colon \R^3 \to \R^4.
    \]
\end{enumerate}
\end{solution}


\newpage
\begin{problem}
Solve the following equation.
\[
\begin{pmatrix} 1 & 1 & 1 \\ 1 & 2 & 1 \\ 1 & 2 & 2 \end{pmatrix} \begin{pmatrix} x \\ y \\ z \end{pmatrix} = \begin{pmatrix} 6 \\ 8 \\ 11 \end{pmatrix}.
\]
\end{problem}
\begin{solution}
First, we create the augmented matrix
\[
\left(\begin{tabular}{ccc|c} 1 & 1 & 1 & 6 \\ 1 & 2 & 1 & 8 \\ 1 & 2 & 2 & 11 \end{tabular} \right).
\]
We can subtract R1 from both R2 and R3 to get
\[
\left(\begin{tabular}{ccc|c} 1 & 1 & 1 & 6 \\ 0 & 1 & 0 & 2 \\ 0 & 1 & 1 & 5 \end{tabular} \right).
\]
Then subtract R3 from R1 to get
\[
\left(\begin{tabular}{ccc|c} 1 & 0 & 0 & 1 \\ 0 & 1 & 0 & 2 \\ 0 & 1 & 1 & 5 \end{tabular} \right).
\]
Finally, subtract R2 from R3 to get
\[
\left(\begin{tabular}{ccc|c} 1 & 0 & 0 & 1 \\ 0 & 1 & 0 & 2 \\ 0 & 0 & 1 & 3 \end{tabular} \right).
\]
This yields our result in the right most column in that $x=1$, $y=2$, and $z=3$.
\end{solution}



\end{document}