%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article} %report allows for chapters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{preamble}

\begin{document}

\begin{center}
   \textsc{\large MATH 271, Homework 2}\\
   \textsc{Due September 11$^\textrm{th}$}
\end{center}
\vspace{.5cm}


\begin{problem} 
Consider the following autonomous equation.
    \[
    x'=-x^2.
    \]
    \begin{enumerate}[(a)]
        \item Draw the phase line for this system. What are the equilibrium point(s)? Which equilibria are stable? Which are unstable? Explain.
        \item Find a general solution to the ODE.
        \item Explain how your general solution fits the qualitative behavior expected from the phase line. That is, can you show that limits of your general solution match your qualitative analysis?
        \item Can $x(0)=0$ be an initial condition? Explain. \emph{Hint: your analysis from the phase line may prove to be more useful than the general solution you found.}
    \end{enumerate}
\end{problem}


\begin{problem}
Let $y$ be a function of $x$ and consider the following differential equation.
\[
y' = y\cos(x).
\]
\begin{enumerate}[(a)]
    \item What is the order of this equation? Is the equation separable? Explain.
    \item Plot an approximation of the slope field for this equation using this Desmos link: \url{https://www.desmos.com/calculator/e93gktwtfo}. \emph{Note that you will have to modify the $g(x,y)$ equation in that page.}
    \item Find the general solution to this equation.
    \item Given the initial data $y(0)=1$, find the particular solution.
    \item Plot this function over your slope field. Explain how you could have approximated this solution using just the slope field.
    \item Explain in words what the solution describes if we let $y(x)$ be the position of some object and $x$ represents time.
\end{enumerate}
\end{problem}

\begin{problem}
Consider the differential equation
\[
x'=\frac{x+t}{t}.
\]
\begin{enumerate}[(a)]
    \item Let $f(x,t)=\frac{x+t}{t}$. Show that $f(x,t)=f(\lambda x, \lambda t)$.
    \item Given (a) holds, use the change of variables $u=\frac{x}{t}$ to rewrite the differential equation as a separable equation in terms of $u$.
    \item Find the general solution to the equation and write your solution in terms of the original variables $t$ and $x$.
\end{enumerate}
\end{problem}

\begin{problem}
Find the general solution to the following equation.
\[
tx'+2x=\frac{\sin(t)}{t}.
\]
Show that your solution is correct. (\emph{Hint: can you use an integrating factor?})
\end{problem}


\end{document}